Exploring Rural Shrink Smart Through Guided Discovery Dashboards

### 2022-09-17 SDSS Journal Paper/Paper Draft.Rmd
Updates from revisions

---
title: Exploring Rural Shrink Smart Through Guided Discovery Dashboards
short-title:  
author:
  - first: Denise                 # first name
    first-init: D.              # initial of first name
    last: Bradford                  # last name
    marker: '1'               # identifier for the affiliation
    #email: denise.bradford@huskers.unl.edu    # optional email address
  - first: Susan
    first-init: S.
    last: VanderPlas
    marker: '1'
    footnote-id: 1
    email: svanderplas@unl.edu 
footnote:
  - content: Corresponding author
    type: corresp               # set type for the corresponding author
    id: 1
affiliation:
  - institution: University of Nebraska - Lincoln
    prefix: Department of Statistics        # optional
    country: United States of America
    marker: 1
classoption:
  - letterpaper    # or a4paper, etc.
  - inpress        # or review
# - linenumber     # add line number for submission
preamble: >
  \usepackage{amsfonts,amsmath,amssymb,amsthm}
  \usepackage{booktabs}
  \usepackage{lipsum}
  \newcommand{\db}[1]{{\textcolor{blue}{#1}}}
  \newcommand{\svp}[1]{{\textcolor{red}{#1}}}

abstract: >
  Many small and rural places are shrinking. Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools. In our paper, we will explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions. Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst.
keywords:
# alphabetical; excluding anything in the title already; lower case
  - Interactive Dashboards
  - Exploratory Data Analysis (EDA)
  - Parallel Coordinate Plots (PCP)
  - Guided Discovery Learning (GDL)
bibliography: jdsart-sample.bib
output: 
  jds.rmd::pdf_article:
    keep_tex: true
---

# Abstract

Many small and rural places are shrinking. Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools. In our paper, we will explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions. Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst.

@@

Data collected from the Index of Relative Rurality (IRR) [@Rural_classification]  were used in the SCC dashboard to help classify the towns. The Index of Relative Rurality (IRR) is a continuous, threshold-free, and unit-free measure of rurality. It is an alternative to the traditional discrete threshold-based classifications.The IRR ranges between 0 (low level of rurality, i.e., urban) and 1 (most rural). Four steps are involved in its design:
\begin{enumerate}
\item Identifying the dimensions of rurality: population size, density, remoteness, and built-up area.
\item Selecting measurable variables to adequately represent each dimension:
    \begin{itemize}
        \item Size: logarithm of population size
        \item Density: logarithm of population density.
        \item Remoteness: network distance.
        \item Built-up area: urban area (as defined by the US Census Bureau) as a percentage of total land area.
    \end{itemize}
\item Re-scaling the variables onto bounded scales that range from 0 to 1.
\item Selecting a link function: unweighted average of the four re-scaled variable.
\end{enumerate}

@@

One factor at the forefront of our visualization design is the importance of reducing the cognitive demands on viewers: we have assembled an incredible amount of data, and it is easy for even statisticians who deal with much larger datasets to get lost in the details of this data. At the same time, we want to invite viewers to engage with the data - to imagine, to draw comparisons, to generalize across towns, and to integrate outside information into the conclusions drawn based on the data we present.
This invitation to engage with the data is similar to the approach advocated in Guided Discovery Learning, a framework leverages hints, feedback, and other helpful information to guide users in interactive exploration [@dedonno].

@@

As a consequence of this massively multivariate structure, we very quickly focused on the use of parallel coordinate plots; other alternatives, such as tours [@tourr], require much more sustained attention to interactive plots as well as a deeper understanding of projections in multidimensional space which we cannot assume our users will have. Introduced in the 1880s [@dOcagne:1885], parallel coordinate or parallel set plots feature a series of vertical axes representing different variables arranged horizontally, with lines connecting each observation. When representing categorical data, parallel set plots may show "blocks" of data instead of individual lines, and are useful for representing conditional relationships between adjacent variables [@Bendix:2005]; modifications of this design, such as common-angle plots [@Hofmann:2013], address the issues which arise due to line-width illusions [@sine, @sineillusion]. Parallel coordinate plots have been generalized to allow for continuous data and additional summaries beyond individual data points, such as densities [@density-pcp]. In this paper, we use the \texttt{ggpcp} package, which leverages the grammar-of-graphics framework introduced in @ggplot2, allowing us to use not only parallel coordinate plots, but also to overlay other statistical summaries, such as boxplots or violin plots, to provide additional context about the marginal distributions of each variable in addition to allowing for exploration of the multivariate space.

@@

Taken together, the app is a single page; the initial "solid ground" which the user explores from consists of maps showing the route from the center of town to necessary services, including the fire department, schools, post offices, and hospitals. In version 2, as shown in \autoref{fig:v2}, the map portion is condensed, and more space is given to value boxes that show vital statistics about the town's QoL and financial metrics. This relatively straightforward display is followed by a parallel coordinate plot that allows the user to see similar towns along dimensions such as economic indicators or population size.

@@

- Our x-axis had a large number of variables that we as researchers believed to be the most strongly associated with quality of life. However, there were still too many variables for users to successfully parse. We reduced the number of variables, focusing on variables that had the highest data quality, and we grouped these variables by quality of life factor [@petersCommunityResiliencyDeclining2019]. 
- Originally, parallel coordinate bands were scaled based on the selected comparison towns. This had the effect of truncating the range of variables and over-emphasizing differences between selected towns relative to the overall range of each variable over all towns in our data set. We chose to show all towns in the data set in a very light $\alpha$ grey color to provide some information about the overall range of each variable. Unfortunately, even with the low-$\alpha$ value, this increased the visual complexity of the plot and confused users. Future iterations will likely make use of another aesthetic, such as boxplots or violin plots, to show the range of values for all towns, and then use lines only for towns that are selected by the user. This should strike a balance between visual complexity and representing the data accurately.
- We noticed that users did not make use of our suggested comparison towns, and so we removed that option in favor of allowing users to enter their own comparison towns directly. Users already had pre-determined towns they wanted to compare to, and our suggestions were just in the way.

@@

We also recognize that we need to leverage the expertise of others in our research team: we are working with artists, researchers in architecture, economists, and sociologists; these researchers provide outside knowledge that we do not have and may be able to help us create insightful use-cases to showcase the app and teach towns how to use it. We can also leverage the app to connect users with our research team, providing additional value to those who use the applet and facilitating development of strategies for maintaining quality of life amid shrinking populations.
<!-- While the idea to use the guided discovery methodology created a lack of understanding in the plot for a novice user on the PCP. The need to explain a visual means that the user is more likely to not use the plot effectively. Our hope was that the PCP would be a bit more "self-explanatory" than it was initially. The number of helpful hints clutters the UI and may be a distraction from the data and the usefulness of the PCPs.  -->

@@

EDA can be a difficult for a variety of groups of people, novice users and experienced researchers. One of the more difficult components of this project has been clearly articulating the purposes of EDA to a diverse group of researchers unfamiliar with the concept. One of the most useful parts of this dashboard iteration process has been as an aid to data discovery: that is, the dashboard motivated us to find additional data sources and incorporate them into the project. Having conversations with other researchers about the EDA process helped to facilitate these conversations, as each discussion seemed to uncover additional data sources that someone remembered after looking at the dashboard. While this facet of the dashboard process may be difficult to study formally, it would be an interesting avenue for investigation.

>>>>

---
title: Exploring Rural Shrink Smart Through Guided Discovery Dashboards
short-title:  
author:
  - first: Denise                 # first name
    first-init: D.              # initial of first name
    last: Bradford                  # last name
    marker: '1'               # identifier for the affiliation
    #email: denise.bradford@huskers.unl.edu    # optional email address
  - first: Susan
    first-init: S.
    last: VanderPlas
    marker: '1'
    footnote-id: 1
    email: svanderplas@unl.edu 
footnote:
  - content: Corresponding author
    type: corresp               # set type for the corresponding author
    id: 1
affiliation:
  - institution: University of Nebraska - Lincoln
    prefix: Department of Statistics        # optional
    country: United States of America
    marker: 1
classoption:
  - letterpaper    # or a4paper, etc.
  - inpress        # or review
# - linenumber     # add line number for submission
preamble: >
  \usepackage{amsfonts,amsmath,amssymb,amsthm,ulem}
  \usepackage{booktabs}
  \usepackage{lipsum}
  \newcommand{\db}[1]{{\textcolor{green}{#1}}}
  \newcommand{\svp}[1]{{\textcolor{red}{#1}}}

abstract: >
  Many small and rural places are shrinking. Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools. In our paper, we will \db{use Iowa data to } explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions. Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst.
keywords:
# alphabetical; excluding anything in the title already; lower case
  - Interactive Dashboards
  - Exploratory Data Analysis (EDA)
  - Parallel Coordinate Plots (PCP)
  - Guided Discovery Learning (GDL)
bibliography: jdsart-sample.bib
output: 
  jds.rmd::pdf_article:
    keep_tex: true
---

\db{\sout{Abstract}}

\db{\sout{Many small and rural places are shrinking. Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools. In our paper, we will explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions. Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst. }}

@@

Data collected from the Index of Relative Rurality (IRR) [@Rural_classification]  were used in the SCC dashboard to help classify the towns. The \db{\sout{Index of Relative Rurality (}}IRR\db{\sout{)}} is a continuous, threshold-free, and unit-free measure of rurality. It is an alternative to the traditional discrete threshold-based classifications.The IRR ranges between 0 (low level of rurality, i.e., urban) and 1 (most rural). Four steps are involved in its design:
\begin{enumerate}
\item Identifying the dimensions of rurality: population size, density, remoteness, and built-up area.
\item Selecting measurable variables to adequately represent each dimension:
    \begin{itemize}
        \item Size: logarithm of population size
        \item Density: logarithm of population density.
        \item Remoteness: network distance.
        \item Built-up area: urban area (as defined by the US Census Bureau) as a percentage of total land area.
    \end{itemize}
\item Re-scaling the variables onto bounded scales that range from 0 to 1.
\item Selecting a link function: unweighted average of the four re-scaled variable.
\end{enumerate}

@@

One factor at the forefront of our visualization design is the importance of reducing the cognitive demands on viewers: we have assembled an incredible amount of data, and it is easy for even statisticians who deal with much larger datasets to get lost in the details of this data. At the same time, we want to invite viewers to engage with the data - to imagine, to draw comparisons, to generalize across towns, and to integrate outside information into the conclusions drawn based on the data we present.
This invitation to engage with the data is similar to the approach advocated in Guided Discovery Learning, a framework \db{that} leverages hints, feedback, and other helpful information to guide users in interactive exploration [@dedonno].

@@

As a consequence of this massively multivariate structure, we very quickly focused on the use of parallel coordinate plots; other alternatives, such as tours [@tourr], require much more sustained attention to interactive plots as well as a deeper understanding of projections in multidimensional space which we cannot assume our users will have. Introduced in the 1880s [@dOcagne:1885], parallel coordinate or parallel set plots feature a series of vertical axes representing different variables arranged horizontally, with lines connecting each observation. When representing categorical data, parallel set plots may show "blocks" of data instead of individual lines, and are useful for representing conditional relationships between adjacent variables [@Bendix:2005]; modifications of this design, such as common-angle plots [@Hofmann:2013], address the issues which arise due to line-width illusions [@sine, @sineillusion]. Parallel coordinate plots have been generalized to allow for continuous data and additional summaries beyond individual data points, such as densities [@density-pcp]. In this paper, we use the \texttt{ggpcp} package \db{in R [@Rlanguage]}, which leverages the grammar-of-graphics framework introduced in @ggplot2, allowing us to use not only parallel coordinate plots, but also to overlay other statistical summaries, such as boxplots or violin plots, to provide additional context about the marginal distributions of each variable in addition to allowing for exploration of the multivariate space.

@@

Taken together, the app is a single page; the initial "solid ground" which the user explores from consists of maps showing the route from the center of town to necessary services, including the fire department, schools, post offices, and hospitals. In version 2, as shown in \autoref{fig:v2}, the map portion is condensed, and more space is given to value boxes that show vital statistics about the town's \db{Quality of Life (QoL)} and financial metrics. This relatively straightforward display is followed by a parallel coordinate plot that allows the user to see similar towns along dimensions such as economic indicators or population size.

@@

- Our x-axis had a large number of variables that we as researchers believed to be the most strongly associated with \db{QoL, grouped into four domains: physical, psychological, social, and environmental}. However, there were still too many variables for users to successfully parse. We reduced the number of variables, focusing on variables that had the highest data quality, and we grouped these variables by \db{\sout{quality of life} QoL} factor [@petersCommunityResiliencyDeclining2019]. 
- Originally, parallel coordinate bands were scaled based on the selected comparison towns. This had the effect of truncating the range of variables and over-emphasizing differences between selected towns relative to the overall range of each variable over all towns in our data set. We chose to show all towns in the data set in a very light $\alpha$ grey color to provide some information about the overall range of each variable. Unfortunately, even with the low-$\alpha$ value, this increased the visual complexity of the plot and confused users. Future iterations will likely make use of another aesthetic, such as boxplots or violin plots, to show the range of values for all towns, and then use lines only for towns that are selected by the user. This should strike a balance between visual complexity and representing the data accurately.
- We noticed that users did not make use of our suggested comparison towns, and so we removed that option in favor of allowing users to enter their own comparison towns directly. Users already had pre-determined towns they wanted to compare to, and our suggestions were just in the way.

@@

We also recognize that we need to leverage the expertise of others in our research team: we are working with artists, researchers in architecture, economists, and sociologists; these researchers provide outside knowledge that we do not have and may be able to help us create insightful use-cases to showcase the \db{applet} and teach towns how to use it. We can also leverage the app to connect users with our research team, providing additional value to those who use the applet and facilitating development of strategies for maintaining \db{\sout{quality of life} QoL} amid shrinking populations.
<!-- While the idea to use the guided discovery methodology created a lack of understanding in the plot for a novice user on the PCP. The need to explain a visual means that the user is more likely to not use the plot effectively. Our hope was that the PCP would be a bit more "self-explanatory" than it was initially. The number of helpful hints clutters the UI and may be a distraction from the data and the usefulness of the PCPs.  -->

@@

EDA can be \db{\sout{a}} difficult for a variety of groups of people, novice users and experienced researchers. One of the more difficult components of this project has been clearly articulating the purposes of EDA to a diverse group of researchers unfamiliar with the concept. One of the most useful parts of this dashboard iteration process has been as an aid to data discovery: that is, the dashboard motivated us to find additional data sources and incorporate them into the project. Having conversations with other researchers about the EDA process helped to facilitate these conversations, as each discussion seemed to uncover additional data sources that someone remembered after looking at the dashboard. While this facet of the dashboard process may be difficult to study formally, it would be an interesting avenue for investigation.

### 2022-09-17 SDSS Journal Paper/Paper-Draft.tex
Updates from revisions

\usepackage{amsfonts,amsmath,amssymb,amsthm} \usepackage{booktabs} \usepackage{lipsum} \newcommand{\db}[1]{{\textcolor{blue}{#1}}} \newcommand{\svp}[1]{{\textcolor{red}{#1}}}

@@

\begin{abstract}
Many small and rural places are shrinking. Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools. In our paper, we will explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions. Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst.
\end{abstract}

@@

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

Many small and rural places are shrinking. Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools. In our paper, we will explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions. Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst.

@@

Data collected from the Index of Relative Rurality (IRR) \citep{Rural_classification} were used in the SCC dashboard to help classify the towns. The Index of Relative Rurality (IRR) is a continuous, threshold-free, and unit-free measure of rurality. It is an alternative to the traditional discrete threshold-based classifications.The IRR ranges between 0 (low level of rurality, i.e., urban) and 1 (most rural). Four steps are involved in its design:

@@

One factor at the forefront of our visualization design is the importance of reducing the cognitive demands on viewers: we have assembled an incredible amount of data, and it is easy for even statisticians who deal with much larger datasets to get lost in the details of this data. At the same time, we want to invite viewers to engage with the data - to imagine, to draw comparisons, to generalize across towns, and to integrate outside information into the conclusions drawn based on the data we present.
This invitation to engage with the data is similar to the approach advocated in Guided Discovery Learning, a framework leverages hints, feedback, and other helpful information to guide users in interactive exploration \citep{dedonno}.

@@

As a consequence of this massively multivariate structure, we very quickly focused on the use of parallel coordinate plots; other alternatives, such as tours \citep{tourr}, require much more sustained attention to interactive plots as well as a deeper understanding of projections in multidimensional space which we cannot assume our users will have. Introduced in the 1880s \citep{dOcagne:1885}, parallel coordinate or parallel set plots feature a series of vertical axes representing different variables arranged horizontally, with lines connecting each observation. When representing categorical data, parallel set plots may show ``blocks'' of data instead of individual lines, and are useful for representing conditional relationships between adjacent variables \citep{Bendix:2005}; modifications of this design, such as common-angle plots \citep{Hofmann:2013}, address the issues which arise due to line-width illusions \citep[\citet{sineillusion}]{sine}. Parallel coordinate plots have been generalized to allow for continuous data and additional summaries beyond individual data points, such as densities \citep{density-pcp}. In this paper, we use the \texttt{ggpcp} package, which leverages the grammar-of-graphics framework introduced in \citet{ggplot2}, allowing us to use not only parallel coordinate plots, but also to overlay other statistical summaries, such as boxplots or violin plots, to provide additional context about the marginal distributions of each variable in addition to allowing for exploration of the multivariate space.

@@

Taken together, the app is a single page; the initial ``solid ground'' which the user explores from consists of maps showing the route from the center of town to necessary services, including the fire department, schools, post offices, and hospitals. In version 2, as shown in \autoref{fig:v2}, the map portion is condensed, and more space is given to value boxes that show vital statistics about the town's QoL and financial metrics. This relatively straightforward display is followed by a parallel coordinate plot that allows the user to see similar towns along dimensions such as economic indicators or population size.

@@

\begin{itemize}
\tightlist
\item
  Our x-axis had a large number of variables that we as researchers believed to be the most strongly associated with quality of life. However, there were still too many variables for users to successfully parse. We reduced the number of variables, focusing on variables that had the highest data quality, and we grouped these variables by quality of life factor \citep{petersCommunityResiliencyDeclining2019}.
\item
  Originally, parallel coordinate bands were scaled based on the selected comparison towns. This had the effect of truncating the range of variables and over-emphasizing differences between selected towns relative to the overall range of each variable over all towns in our data set. We chose to show all towns in the data set in a very light \(\alpha\) grey color to provide some information about the overall range of each variable. Unfortunately, even with the low-\(\alpha\) value, this increased the visual complexity of the plot and confused users. Future iterations will likely make use of another aesthetic, such as boxplots or violin plots, to show the range of values for all towns, and then use lines only for towns that are selected by the user. This should strike a balance between visual complexity and representing the data accurately.
\item
  We noticed that users did not make use of our suggested comparison towns, and so we removed that option in favor of allowing users to enter their own comparison towns directly. Users already had pre-determined towns they wanted to compare to, and our suggestions were just in the way.
\end{itemize}

@@

Our dashboard design philosophy worked primarily to promote a town-centric approach application with comparisons to other similar towns being secondary. This approach created a way for the user to see their town information at the top of the page and to explore the PCP after reviewing their own town's essential statistics. The PCP in the lower part of the dashboard allowed for the user to see the plot and adjust to the fact that they could add more towns to the plot, providing an opportunity to explore the wider dataset from a base of familiar knowledge.
{\textcolor{red}{While we initially framed the design around guided discovery learning, the approach did not seem to suffice for our user base; instead, we found that users were more drawn to the unfamiliar from the start. We will likely leverage this in future iterations by using visual forms such as flower plots to draw the users in; even though these plots are not ideal for numerical display of data, the visual novelty and aesthetic appeal will provide some motivation to continue exploring and thinking about the data.}}

@@

We also recognize that we need to leverage the expertise of others in our research team: we are working with artists, researchers in architecture, economists, and sociologists; these researchers provide outside knowledge that we do not have and may be able to help us create insightful use-cases to showcase the app and teach towns how to use it. We can also leverage the app to connect users with our research team, providing additional value to those who use the applet and facilitating development of strategies for maintaining quality of life amid shrinking populations.

@@

EDA can be a difficult for a variety of groups of people, novice users and experienced researchers. One of the more difficult components of this project has been clearly articulating the purposes of EDA to a diverse group of researchers unfamiliar with the concept. One of the most useful parts of this dashboard iteration process has been as an aid to data discovery: that is, the dashboard motivated us to find additional data sources and incorporate them into the project. Having conversations with other researchers about the EDA process helped to facilitate these conversations, as each discussion seemed to uncover additional data sources that someone remembered after looking at the dashboard. While this facet of the dashboard process may be difficult to study formally, it would be an interesting avenue for investigation.

>>>>

\usepackage{amsfonts,amsmath,amssymb,amsthm,ulem} \usepackage{booktabs} \usepackage{lipsum} \newcommand{\db}[1]{{\textcolor{green}{#1}}} \newcommand{\svp}[1]{{\textcolor{red}{#1}}}

@@

\begin{abstract}
Many small and rural places are shrinking. Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools. In our paper, we will \db{use Iowa data to } explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions. Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst.
\end{abstract}

@@

{\textcolor{green}{\sout{Abstract}}}

{\textcolor{green}{\sout{Many small and rural places are shrinking. Interactive dashboards are the most common use cases for data visualization and context for exploratory data tools. In our paper, we will explore the specific scope of how dashboards are used in small and rural area to empower novice analysts to make data-driven decisions. Our framework will suggest a number of research directions to better support small and rural places from shrinking using an interactive dashboard design, implementation and use for the every day analyst. }}}

@@

Data collected from the Index of Relative Rurality (IRR) \citep{Rural_classification} were used in the SCC dashboard to help classify the towns. The {\textcolor{green}{\sout{Index of Relative Rurality (}}}IRR{\textcolor{green}{\sout{)}}} is a continuous, threshold-free, and unit-free measure of rurality. It is an alternative to the traditional discrete threshold-based classifications.The IRR ranges between 0 (low level of rurality, i.e., urban) and 1 (most rural). Four steps are involved in its design:

@@

One factor at the forefront of our visualization design is the importance of reducing the cognitive demands on viewers: we have assembled an incredible amount of data, and it is easy for even statisticians who deal with much larger datasets to get lost in the details of this data. At the same time, we want to invite viewers to engage with the data - to imagine, to draw comparisons, to generalize across towns, and to integrate outside information into the conclusions drawn based on the data we present.
This invitation to engage with the data is similar to the approach advocated in Guided Discovery Learning, a framework {\textcolor{green}{that}} leverages hints, feedback, and other helpful information to guide users in interactive exploration \citep{dedonno}.

@@

As a consequence of this massively multivariate structure, we very quickly focused on the use of parallel coordinate plots; other alternatives, such as tours \citep{tourr}, require much more sustained attention to interactive plots as well as a deeper understanding of projections in multidimensional space which we cannot assume our users will have. Introduced in the 1880s \citep{dOcagne:1885}, parallel coordinate or parallel set plots feature a series of vertical axes representing different variables arranged horizontally, with lines connecting each observation. When representing categorical data, parallel set plots may show ``blocks'' of data instead of individual lines, and are useful for representing conditional relationships between adjacent variables \citep{Bendix:2005}; modifications of this design, such as common-angle plots \citep{Hofmann:2013}, address the issues which arise due to line-width illusions \citep[\citet{sineillusion}]{sine}. Parallel coordinate plots have been generalized to allow for continuous data and additional summaries beyond individual data points, such as densities \citep{density-pcp}. In this paper, we use the \texttt{ggpcp} package {\textcolor{green}{in R [@Rlanguage]}}, which leverages the grammar-of-graphics framework introduced in \citet{ggplot2}, allowing us to use not only parallel coordinate plots, but also to overlay other statistical summaries, such as boxplots or violin plots, to provide additional context about the marginal distributions of each variable in addition to allowing for exploration of the multivariate space.

@@

Taken together, the app is a single page; the initial ``solid ground'' which the user explores from consists of maps showing the route from the center of town to necessary services, including the fire department, schools, post offices, and hospitals. In version 2, as shown in \autoref{fig:v2}, the map portion is condensed, and more space is given to value boxes that show vital statistics about the town's {\textcolor{green}{Quality of Life (QoL)}} and financial metrics. This relatively straightforward display is followed by a parallel coordinate plot that allows the user to see similar towns along dimensions such as economic indicators or population size.

@@

\begin{itemize}
\tightlist
\item
  Our x-axis had a large number of variables that we as researchers believed to be the most strongly associated with {\textcolor{green}{QoL, grouped into four domains: physical, psychological, social, and environmental}}. However, there were still too many variables for users to successfully parse. We reduced the number of variables, focusing on variables that had the highest data quality, and we grouped these variables by {\textcolor{green}{\sout{quality of life} QoL}} factor \citep{petersCommunityResiliencyDeclining2019}.
\item
  Originally, parallel coordinate bands were scaled based on the selected comparison towns. This had the effect of truncating the range of variables and over-emphasizing differences between selected towns relative to the overall range of each variable over all towns in our data set. We chose to show all towns in the data set in a very light \(\alpha\) grey color to provide some information about the overall range of each variable. Unfortunately, even with the low-\(\alpha\) value, this increased the visual complexity of the plot and confused users. Future iterations will likely make use of another aesthetic, such as boxplots or violin plots, to show the range of values for all towns, and then use lines only for towns that are selected by the user. This should strike a balance between visual complexity and representing the data accurately.
\item
  We noticed that users did not make use of our suggested comparison towns, and so we removed that option in favor of allowing users to enter their own comparison towns directly. Users already had pre-determined towns they wanted to compare to, and our suggestions were just in the way.
\end{itemize}

@@

Our dashboard design philosophy worked primarily to promote a town-centric approach application with comparisons to other similar towns being secondary. This approach created a way for the user to see their town information at the top of the page and to explore the PCP after reviewing their own town's essential statistics. The PCP in the lower part of the dashboard allowed for the user to see the plot and adjust to the fact that they could add more towns to the plot, providing an opportunity to explore the wider dataset from a base of familiar knowledge.

While we initially framed the design around guided discovery learning, the approach did not seem to suffice for our user base; instead, we found that users were more drawn to the unfamiliar from the start. We will likely leverage this in future iterations by using visual forms such as flower plots to draw the users in; even though these plots are not ideal for numerical display of data, the visual novelty and aesthetic appeal will provide some motivation to continue exploring and thinking about the data.

@@

We also recognize that we need to leverage the expertise of others in our research team: we are working with artists, researchers in architecture, economists, and sociologists; these researchers provide outside knowledge that we do not have and may be able to help us create insightful use-cases to showcase the {\textcolor{green}{applet}} and teach towns how to use it. We can also leverage the app to connect users with our research team, providing additional value to those who use the applet and facilitating development of strategies for maintaining {\textcolor{green}{\sout{quality of life} QoL}} amid shrinking populations.

@@

EDA can be {\textcolor{green}{\sout{a}}} difficult for a variety of groups of people, novice users and experienced researchers. One of the more difficult components of this project has been clearly articulating the purposes of EDA to a diverse group of researchers unfamiliar with the concept. One of the most useful parts of this dashboard iteration process has been as an aid to data discovery: that is, the dashboard motivated us to find additional data sources and incorporate them into the project. Having conversations with other researchers about the EDA process helped to facilitate these conversations, as each discussion seemed to uncover additional data sources that someone remembered after looking at the dashboard. While this facet of the dashboard process may be difficult to study formally, it would be an interesting avenue for investigation.
